Finding and controling biases in dark matter halo concentrations estimates

### 2016-07-04 paper/ms.tex
new title new perspective
new plot in ms
new abstract

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{A new method to estimate dark matter halo concentrations}
\author{
  C. N. Poveda-Ruiz \thanks{cn.poveda542@uniandes.edu.co}$^{1}$
  J. E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{1}$
  J. C. Mu\~noz-Cuartas \thanks{juan.munozc@udea.edu.co}$^{2}$
}

@@

\begin{abstract}
We present a new algorithm to estimate the concentration of N-body
dark matter halos using the integrated mass profile.
The method uses the full particle information without any binning,
making it reliable in cases when low numerical resolution becomes a
limitation for other methods.   
We compare the new algorithm against two other methods:
maximum radial velocity measurements and radial particle binning.
Tests on mock and N-body halos show that the accuracy of the new method to
measure concentrations varies with halo resolution, outperforming the
other two methods.  
We also esteimate the mass-concentration relationship on N-body data. 
We find that in the probed mass range ($10^{10}\hMsun < M_{h} <
10^{14}\hMsun$) the three methods give consistent results within the
statistical uncertainties.   
There is a small deviation at low masses, $M<10^{12}\hMsun$, where
the new method yields lower median concentration values by $25\%$
compared to the velocity method; the difference is explained by the bias 
that the velocity method introduces.
From these results we believe that this new method is a promising tool
to probe the internal structure of dark matter halos.
\end{abstract}

@@

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.49\textwidth]{avg_foff_bolshoi.pdf}
\end{center}
\vspace{-0.5cm}
\caption{Average value of the fractional offset ($f_{\rm
    off}$Eq. \ref{eq:f_off}) between the concentration in a halo at
  high resolution versus its downsampled versions.  Each point is the
  result of 100 downsampling iterations.  The velocity method
  noticeably overestimates the concentration up to a factor of $0.25$,
  while the new method only underestimates the concentrations by up to
  a factor of $0.05$.
    \label{fig:downsampling}}
\end{figure}

>>>>

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{Finding and controling biases in dark matter halo concentrations estimates}
\author{
  C. N. Poveda-Ruiz \thanks{cn.poveda542@uniandes.edu.co}$^{1}$
  J. E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{1}$
  J. C. Mu\~noz-Cuartas \thanks{juan.munozc@udea.edu.co}$^{2}$
}

@@

\begin{abstract}
We present a jacknife method to estimate the bias on the estimated
concentration values as a function of particle number in the halo. 
We find that algorithms based on the maximum radial velocity
and radial particle binning show similar biases,
overestimating concentrations by a factor of $15\%-20\%$ for
halos sampled with $200$ particles.  
To overcome this bias we propose a new concentration estimation
algorithm that uses the integrated mass profile.
The method uses the full particle information without any binning,
making it reliable in cases when low numerical resolution becomes a
limitation for other methods, reducing the bias to $2\%-3\%$ for halos
sampled with $200$ particles. 
To keep the biases of the velocity and density methods to the same
low levels, dark matter halos should be sampled with at least $\sim
1000$ particles.
We also find that the mass-concentration relationship could be
shallower than expected from numerical experiments that do not take
into account this bias. 
From these results we believe that jacknife downsampling and the
concentration estimate based on the integral mass densities are
promising tools to probe the internal structure of dark matter halos
in numerical simulations.
\end{abstract}

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.48\textwidth]{avg_foff_bolshoi.pdf}
  \includegraphics[width=0.48\textwidth]{avg_foff_vialactea.pdf}
\end{center}
\vspace{-0.5cm}
\caption{Average value of the fractional offset ($f_{\rm
    off}$Eq. \ref{eq:f_off}) between the concentration in a halo at
  high resolution versus its downsampled versions.  Each point is the
  result of 100 downsampling iterations.  The velocity method
  noticeably overestimates the concentration up to a factor of $0.25$,
  while the new method only underestimates the concentrations by up to
  a factor of $0.05$.
    \label{fig:downsampling}}
\end{figure*}

@@
